\documentclass[11pt,reqno]{amsart}
\usepackage[utf8]{inputenc}
\usepackage{braket,amsmath,amsfonts,amsthm,url,graphicx,url,xcolor,hyperref,subcaption, multicol,dsfont, comment}



\oddsidemargin-0cm
\evensidemargin-0cm
\textwidth16.1cm
%\topmargin1cm
%\footskip1.3cm %1.5cm
%\footheight3cm
%\textheight21.5cm %23cm
%\parindent0em
\renewcommand{\baselinestretch}{1.1}

\setlength{\parindent}{4ex}

\newtheorem{lemma}{Lemma}[section]
\newtheorem{prop}[lemma]{Proposition}
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{thmx}{Theorem}
\newtheorem{cor}[lemma]{Corollary}
\newtheorem{con}[lemma]{Conjecture}
\newtheorem{prob}[lemma]{Problem}
\newtheorem{rem}[lemma]{Remark}
\newtheorem{definition}[lemma]{Definition}
\newtheorem{frem}[lemma]{Final remark}
\newtheorem{fact}[lemma]{Fact}
\newtheorem{claim}[lemma]{Claim}
\newtheorem{assump}[lemma]{Assumption}
\newtheorem{exam}[lemma]{Example}
\newtheorem{cexam}[lemma]{Counterexample}
\newtheorem{quest}[lemma]{Question}


\newcommand{\expec}{\mathbb{E}}
\newcommand{\Hil}{\mathcal{H}}
\newcommand{\ketbra}[1]{|{#1}\rangle\langle{#1}|}


\newcommand{\Haar}{H}

\newcommand{\NN}{\mathbb{N}}
\newcommand{\MM}{\mathbb{M}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\BB}{\mathbb{B}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\E}{\mathcal{E}}
\renewcommand{\S}{\mathcal{S}}
\newcommand{\sep}{\square}
\newcommand{\id}{\hat{1}}
\newcommand{\Id}{\mathrm{Id}}

\newcommand{\cV}{\mathcal{V}}
\newcommand{\cW}{\mathcal{W}}
\newcommand{\one}{\mathds{1}}

\newcommand{\pl}{\hspace{.1cm}}

\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\N}{\mathcal{N}}
\newcommand{\R}{\mathcal{R}}
\newcommand{\G}{\mathcal{G}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\idd}{\text{id}}
\newcommand{\Rot}{\text{Rot}}
\newcommand{\Ha}{H}
\newcommand{\tr}{\text{tr}}
\newcommand{\epow}[2]{F^{({#2})}_{{#1}}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\cT}{\mathcal{T}}

\newcommand{\Dmax}{D_{\infty}}
\newcommand{\BP}{\text{BP}}
\newcommand{\polylog}{\text{polylog}}

\newcommand{\newterm}[1]{{\textit{#1}}}

\newcommand{\precsucc}[2]{\text{ } {}^{\prec {#2}}_{\succ {#1}} \text{ } }

\newcommand{\compl}[2]{\C({#1} \subseteq {#2})}

\renewcommand{\deg}[1]{\mathrm{deg}({#1})}

\DeclareUnicodeCharacter{2009}{\,} 

\title{Notes on Designs from Lindbladians}
%Shallow, Unstructured, Approximate k-Designs via Quantum Relative Entropy


%\author[N. Laracuente]{Nicholas Laracuente}
%\address{Indiana University Bloomington. Bloomington, IN. 47404} \email[Nicholas Laracuente]{nlaracu@iu.edu}

\thanks{\hspace{-5mm} Part of this work was completed as an IBM Postdoc at the University of Chicago.}



\newcommand{\mbalign}{}
\newcommand{\mbnline}{}

\newcommand{\todo}[1]{{\color{red}\textbf{TODO:} #1}}

\begin{document}

\section{Background and Previous Results}
We define the $k$-fold twirl weighted by probability measure $\mu$ on $U(d)$ to be
\begin{equation}
    \Phi_{\mu, k} := \int U^{\otimes k} \rho U^{\otimes k \dagger} d \mu(U)
\end{equation}
on an input system of dimension $d^k$.
\begin{lemma}[\cite{laracuente_quantum_2025}, Lemma 3.1] \label{lem:chainmin1}
Let $(\Phi_j)_{j=1}^m$ be a given family of quantum channels with respective $(\lambda_j)_{j=1}^m$-SDPI and decoherence-free subspace projections $(\E_j)_{j=1}^m$, $\E$ a joint decoherence-free subspace projection commuting with each $\Phi_j$ (for which $\E \E_j = \E_j \E = \E$),
%\fl{The 'joint' here seems to refer to the fact that $\E_m\circ \E = \E$? You seem to use this when applying the chain rule in \eqref{eq:chain2} below.}
and $(\Psi_j)_{j=1}^m$ a family of channels each commuting with $\E$. Then
%\fl{What are the `rotations' that you're referring to here? Do you mean the $\E_j$?}
\begin{equation*}
    \begin{split}
         D(\rho \| \E(\rho)) - D((\Psi_m \Phi_m) \dots (\Psi_1 \Phi_1) (\rho) \| \E (''))
            \mbnline \mbalign \geq \sum_{j=1}^m \lambda_j D( (\Psi_{j-1} \Phi_{j-1}) \dots (\Psi_1 \Phi_1)(\rho) \| \E_j ('')) \pl,
    \end{split}
\end{equation*}
where when $''$ appears in the second argument to relative entropy, it is equal to the first argument.
\end{lemma}
\begin{prop}[\cite{laracuente_quasi-factorization_2022}, Proposition I.6] \label{prop:decay-merge}
Let $\{\Phi^t_j : j \in 1...J \in \NN\}$ be GNS self-adjoint quantum Markov semigroups such that $\Phi_j^t = \exp(- \L_j t)$ with fixed point conditional expectation $\E_{j} = \lim_{t \rightarrow \infty} \Phi_j^t$ admitting common invariant state $\sigma$. Let $\E_{\sigma *}$ be the weighted intersection fixed point conditional expectation. Let $\Phi^t$ be the semigroup generated by $\L = \sum_j \alpha_j \L_j + \L_0$, where $\L_0$ generates $\Phi_0^t$ such that $\Phi_0^t \E_{\sigma *} = \E_{\sigma *} \Phi_0^t = \E_{\sigma *}$. If the conditional expectations $(\alpha_j)$-quasi-factorize, and each $\Phi^t_j$ has $\lambda$-(C)MLSI, then $\Phi^t$ has $\lambda$-(C)MLSI.
\end{prop}
\begin{lemma}[Chain Rule] \label{lem:chainexp}
Let $\omega$ be a density and $\E$ be a conditional expectation such that $\E(\omega) = \omega$. Then for any density $\rho$,
\begin{equation*}
D(\rho \| \omega) = D(\rho \| \E(\rho)) + D(\E(\rho) \| \omega) \pl.
\end{equation*}
\end{lemma}
Lemma \ref{lem:chainexp} is a well-known identity in quantum information. An important consequence of Lemma \ref{lem:chainexp} is that for any channel $\Phi$ and conditional expectation $\E'$,
\begin{equation} \label{eq:chain-iter}
    D(\rho \|\Psi(\rho)) + D(\rho \| \E'(\rho))
        \geq D(\E'(\rho) \| \E' \circ \Psi(\rho)) + D(\rho \| \E'(\rho))
    = D(\rho \| \E' \circ \Psi(\rho)) \pl.
\end{equation}
\begin{lemma}[\cite{laracuente_quasi-factorization_2022}, Corollary II.15. \cite{gao_complete_2025}, Lemma 2.3]  \label{lem:revconv}
Let $\rho$ be a density and $\E,\Psi$ be quantum channels such that
\begin{equation}
    (1 - \epsilon) \E \prec \Psi(\rho) \prec (1 + \delta) \E
\end{equation}
%\begin{equation}
%    (1-\epsilon) \E(\rho) + \epsilon \Phi(\rho) \leq (1 + \delta) \E(\rho)
%\end{equation}
for constants $\epsilon, \delta \in (0,1)$. Assume $\rho \in \text{supp}(\E(\rho))$. Furthermore, assume $\Psi \E = \E$. Then
\begin{equation}
D(\rho \| \Psi(\rho)) \geq \beta_{\epsilon, \delta} D(\rho \| \E(\rho))
\end{equation}
with
\begin{equation} \label{eq:simpleG} 
    \beta_{\epsilon, \delta} \geq \frac{1}{{(1 + \epsilon)(1 + \delta)}} \Big ( 1 - \frac{2 (1 + \epsilon) \delta^2}{(\epsilon + \delta)(\ln (1 + \delta / \epsilon) - 1) + \epsilon} - 4 \epsilon - \epsilon^2 \Big ) \pl.
\end{equation}
If $\epsilon = \delta$, then
\begin{equation}
    \beta_{\epsilon, \epsilon} \geq \frac{1-\epsilon}{1 + \epsilon} - \frac{\epsilon}{(1 - \epsilon) (2 \ln 2 - 1)} \pl, \text{ and } \beta_{\epsilon, \epsilon} \geq 1 - 12 \epsilon \pl.
\end{equation}
If $\epsilon \leq 1/10$, then $\beta_{\epsilon, \epsilon} \geq 1/2$.
\end{lemma}
\begin{definition} \label{def:quasifactorization}
    Let $\{\E_j\}_{j=1}^m$ be a set of conditional expectations with the common invariant state $\sigma$ and $\sigma$-weighted intersection subalgebra projection $\E$. We say that the set $(\alpha_j)$-quasi-factorizes (in other texts, has $\alpha$-approximate tensorization) if
    \begin{equation}
        \sum_j \alpha_j D(\cdot \| \E_j(\cdot)) \geq D(\cdot \| \E(\cdot)) \pl.
    \end{equation}
\end{definition}
Inspired by quasi-factorization and by the methods of \cite{schuster_random_2025, laracuente_approximate_2026} is an the following condition:
\begin{assump}[Entropic Gluing] \label{assump:glue} \normalfont
    Let $\E_{A}$ and $\E_B$ be subspace projections and $\E$ project to their intersection subspace. For a situation-dependent $\epsilon > 0$,
    \begin{itemize}
        \item (Immediate) $D(\rho \| \E_{A} \E_B (\rho)) \geq (1-\epsilon) D(\rho \| \E(\rho))$;
        \item ($r$-Repeated) $D(\rho \| (\E_{A}\E_B)^r (\rho)) \geq (1-\epsilon) D(\rho \| \E(\rho))$ for $\rho$-independent $r \in \NN$.
    \end{itemize}
    
\end{assump}
A relative error analog of the gluing assumption was proven to hold for $k$-fold unitary twirls in \cite[Lemma 8]{schuster_random_2025} and a weaker version shown concurrently in \cite{laracuente_approximate_2026}. In that setting of $k$-fold random unitaries, $\E_{A} \circ \E_B \precsucc{5 k^2 / q^m}{} \E$, where $q$ is the local qudit dimension, and $A$ and $B$ are subsystems that overlap on $m$ elementary qudits ($k \times m$ taken $k$-fold). Combined with Lemma \ref{lem:revconv}, the relative error gluing Lemma implies the immediate case of Assumption \ref{assump:glue} with $\epsilon \leq 60 k^2 / q^m$. Assumption \ref{assump:glue} and Equation \eqref{eq:chain-iter} together yield quasi-factorization.

Recall the tightened Fannes-Alicki inequality on relative entropy with respect to a subspace projection \cite[Lemma 7]{winter_tight_2016}:
\begin{lemma}[] \label{lem:wsb}
    if $\rho$ and $\omega$ are states, $\epsilon := \frac{1}{2} \| (\Phi \otimes \Id - \Psi \otimes \Id)(\rho) \|_1$, and $\E$ projects to a convex subspace, then
    \begin{equation} \label{eq:wsb}
        |D(\rho \| \E(\rho)) - D(\omega \| \E(\omega))|
            \leq \epsilon \sup_{\sigma} D(\sigma \| \E(\sigma)) + (1+\epsilon) h \Big ( \frac{\epsilon}{1+\epsilon} \Big ) \pl.
    \end{equation}
\end{lemma}

\section{General: Tree Decomposition}
Consider a Lindbladian $\L = \sum_{j=1}^m \L_j$. The overarching goal of this section is to derive as general as possible a way to obtain a decay constant for $\L$ in terms of (1) the decay constants of its constituents (2) how closely compositions of respective fixed point projections approximate their intersection subspaces. In particular, by recursively applying Assumption \ref{assump:glue}, one may often obtain efficient decay estimates for the whole in terms of sums of their parts.

We decompose $\Phi$ or $\L$ via an ordered tree $T$, where the root of $T$ holds $\Phi$ or $\L$, and each parent node holds the composition of channels or Lindbladians included in its children. The leaves of the tree correspond to smaller sub-Lindbladians for which we need an a priora decay constant estimate. %  for which we have some a priori estimate of decay constants. The tree can be redundant - the children of a node need not hold disjoint subsets or subsequences. For Lindbladians, redundancy is almost trivial via the identity $\L_j = \sum_r \kappa_r \L_j / (\sum_r \kappa_r)$ for any positive family $(\kappa_r)$. For sequential channels, overlaps must be restricted to neighboring siblings such that sequential order is preserved up to the overlapping portion. The ability to insert redundancy is used in \cite[Unpublished Revision]{laracuente_quantum_2025} to increase the overlaps between fixed point conditional expectations. We assume compatible entropy decay in that each $\Phi_j$ or $\L_j$ has a non-zero decay constant, and any subsequence (resp subset sum) has a well-defined decoherence-free subspace projection.
By $\L[x]$ we respectively denote the sub-Lindbladian corresponding to node $x$, by $\E[x]$ its conditional expectation, and by $\lambda[x]$ its decay constant. By $\E[x_1, \dots, x_\ell]$ we denote the intersection subspace projection of all nodes listed (assuming it is indeed well-defined). The key is the following observation:
\begin{lemma} \normalfont \label{lem:lindblad-tree}
    For each node $x \in T$, let $y_1, \dots, y_\ell$ be its children.
    %For each index $i$ in the original subsequence, let $(p_{i,j} : j = 1 \dots \ell)$ be any probability distribution supported on all children in which it appears.
    If the family$(\E[y_j])_{j=1}^\ell$ has $(\alpha_j)_{j=1}^{\ell}$-quasi-factorization, and each $\L[y_j]$ has $\lambda \times \alpha_j$-CMLSI, then $\L[x]$ has $\lambda$-CMLSI.
    %If the immediate case of Assumption \ref{assump:glue} holds for each $\E[y_{j+1}]$ and $\E[y_1, \dots, y_j]$ with contant $\epsilon_j$ for each $j \in 1, \dots, \ell-1$, then $\Phi[x]$ or $\L[x]$ has decay constant at least
    %\begin{equation}
    %    \left ( \prod_{j=1}^{\ell-1} (1 - \epsilon_j) \right )
    %        \times \left ( \min_{i,j} \lambda[y_j] p_{i,j} \right ) \pl.
    %\end{equation}
\end{lemma}
Lemma \ref{lem:linblad-tree} follows almost immediately from \ref{prop:decay-merge}. To use Lemma \ref{lem:lindblad-tree} requires two main criteria:
\begin{itemize}
    \item Quasi-factorization holds between different sub-Lindbladians' fixed point projections.
    \item For the leaf subsequences, there is an a priori decay rate estimate.
\end{itemize}
We may think of Lemma \ref{lem:tree} as reducing the size of a problem efficiently. The cost of converting from spectral gaps on the leaf subsequences and then building up relative entropy decay through the tree is often much smaller than that of converting spectral gaps globally.
\begin{prop}
    Let $\L = \sum_{j=1}^m \L_m$ be a Lindbladian with tree decomposition $T$. Assume each $\L_j$ has decay constant at least $\lambda$.
    %and Assumption \ref{assump:glue} holds with constant $\epsilon_{\ell}$
    For each non-leaf node $x$ at level $\ell$, assume for the children $y_1, \dots, y_{r}$ that $\E[y_1], \dots, \E[y_r]$ $\alpha_\ell$-quasi-factorize. % $\E[y_1, \dots, y_r]$ and $\E[y_{r+1}]$ when $y_1, \dots, y_{r+1}$ are siblings at level $\ell$.
    Then $\L$ has decay constant at least $\lambda \times \prod_{\ell=1}^{h_T} \alpha_{\ell}$, where $h_T$ is the height of the tree.
\end{prop}
\begin{proof}
    Assume that the $\ell$th level of $T$ contains nodes $x_1, \dots, x_{m[\ell]}$. By Lemma \ref{lem:tree}, the next level up will have minimum decay constant at least $\min_{j = 1 \dots m[\ell]} \beta(\epsilon_j)^{d_T} \lambda[x_j]$. Applying iteratively completes the Proposition.
\end{proof}
\begin{theorem} \label{thm:regular-tree}
    Let $\L = \sum_{j=1}^m \L_m$ be a Lindbladian with tree decomposition $T$. Assume that Assumption \ref{assump:glue} holds with constant $c / q^r \geq 2$ when $r$ elementary units are shared between adjacent siblings, and at the $\ell$th level starting from the bottom, at least $\log_q (c) + a + \ell - 1$ elementary units are shared for some constant $a \in \NN$.
    Assume each $\L_j$ (resp $\Phi_j$) has decay constant at least $\lambda$. Then $\L$ (resp $\Phi$) has decay constant at least $(1 - 12 / q^a) \lambda$.
\end{theorem}
\begin{proof}
    ...
\end{proof}

\subsection{Lindbladians for Random Circuits}
Consider a system of $n$ qubits and associated connectivity hypergraph $G$. Let $|G|$ denote the number of edges in $G$. We start with the model
\begin{equation} \label{eq:circuit-hypergraph-lind}
    \L^{(k)}_G = \sum_{e \in G} \frac{|e|}{d_e} \L_e^{(k)},
    \pl \L_e^{(k)}(\rho) = \rho - \int U^{\otimes k} \rho U^{\dagger \otimes k} d \Haar_e(U) \pl,
\end{equation}
where $d \Haar_e$ denotes the \textit{local} Haar measure on unitaries on the qudits indexed by edge $e$ (acting as identity on the rest of the system), $|e|$ is the number of nodes connected by $e$, and $d_e$ is the sum of degrees of nodes connected by $e$. Since each $\L_e^{(k)}$ generates state replacement, convexity of relative entropy implies that it has CMLSI constant 1. If $G$ is a graph, then we may think of each $U_e$ as acting on a 2-qudit subsystem. For example, the Lindbladian analog of 1-D brickwork on $n$ qudits of unspecified local dimension would have $G = \{(1,2), (2,3), \dots, (n-1, n)\}$. However, we are primarily interested in much more connected graphs, for which we will obtain more efficient estimates. \todo{Question for Nima: is this equivalent to the ``Brownian circuit'' model? The answer might give some guidance on how to convert Brownian SYK to a similar form.}

\begin{exam}[Random Circuit Lindbladian on a Graph] \normalfont
Assume that $D \geq 2$, and $n = 2^{x D}$ for some $x \in \N$. Let
\begin{equation}
    \L_{\nu}^{(k)} = \int \L_{G}^{(k)} d\nu(G) \pl,
\end{equation}
where $\nu$ is a probability measure on graphs, and each $G$ in its support is a $D$-dimensional lattice hypercube. This $\L_{\nu}$ describes any graph Lindbladian can be thought of as the convex combination of Lindbladians on lattice hypercubes. For example, this scenario includes the complete graph, as well as any connectivity that is already on a $D$-dimensional lattice.

The lattice hypercube structure enables recursively splitting the graph in half, obtaining subgraphs of size $2^{x D - 1}$, then $2^{x D - 2}$, and so on until reaching size $m$ for some $m \in \NN$ we will set later. Successive splits will iterate through axes, so every $D$ splits, we again obtain (smaller) hypercubes. This way, we also see that at each split, the number of edges crossing the split is divided by 2 (conversely if building up, each join involves doubly many edges as its children). These edges can be assigned somewhat arbitrarily to one side of the split or the other. The full recursive splitting requires exactly $\log_2 (n/m)$ layers of recursion. The face size of an $m$-length hyper-rectangle is at least $m^{(D-1)/D}$. In Theorem \ref{thm:regular-tree}, we use the SHH gluing Lemma \ref{schuster_random_2025} for Assumption \ref{assump:glue} to set $c = 5 k^2$, $q$ to the local qudit dimension, and $a = m^{(D-1)/D} - \log_q(5 k^2)$.

To estimate the decay constants of the $m$-size sublattice Lindbladians, we apply Proposition \ref{prop:decay-merge}. Since each $\L_e$ has $1$-CMLSI, each $m$-size sublattice Lindbladian has decay constant proportional to the smallest value of $2/d_e$ times the complete quasi-factorization constant $\alpha$ (recall Definition \ref{def:quasifactorization}). To calculate $\alpha$, note using Equation \eqref{eq:chain-iter} that
\begin{equation}
    \sum_e D(\rho \| \E_e(\rho)) \geq
        \frac{1}{w} D \bigg ( \rho \bigg \| \bigg (\prod_{e \in G} \E_e 
    \bigg )^w (\rho) \bigg )
\end{equation}
for any $w \in \NN$, where the product is ordered arbitrarily. The composition $(\prod_{e \in G} \E_e )^w$ is itself a composition of local random gates. Currently, we might apply \cite[Theorem ??]{belkin_?} along a Hamiltonian path to obtain $\alpha = 1 / m k C(k)$ for some function $C(k)$. In the qubit case, $C(k) = O(\mathrm{polylog}(k))$. Combined with Equation \eqref{eq:circuit-hypergraph-lind}, we obtain that each $m$-size sub-Lindbladian has $2 / m C(k) d_E$-CMLSI. \todo{Can we eliminate the $2 / d_E$ factor, leading to a uniform bound in the spatial dimension or enhancement with higher connectivity?}

Finally, we should choose a good value of $m$. We need that $a = m^{(D-1)/D} - \log_q(5 k^2) > 0$, and ideally it should be large enough that $q^a$ is small. In contrast, the ``leaf-level'' value of $\lambda$ is divided by $m$, so it should not be too large. Set $m$ to the smallest power of $2^D$ that is at least $(\log_q(5 k^2))^{D/(D-1)} + \log_q 24$, so $q^a \geq 24$. Note that $d_e \leq 2^{D+1}$ on lattices. Again recalling Theorem \ref{thm:regular-tree}, $\L_{\nu}$ has
\begin{equation}
    \frac{1}{k \times C(k)\times \max_D 2^{D+1} ((\log_q(5 k^2))^{D/(D-1)} + \log_q 24) } \text{-CMLSI}
\end{equation}
if ``$\max_D$'' is with respect to all edges in all graphs in $\nu$'s support. The important aspects are that (1) the estimate is independent of $n$ (2) as long as $C(k) = O(\text{polylog(k)})$, $\L_{\nu}$ has $1/k \times O(\mathrm{polylog(k)})$-CMLSI.
\end{exam}

\todo{Next step: study hypergraph decomposition in this model. Can we decompose complete or other hypergraphs into convex combinations of lattice-like objects?}

Idea: \cite{laracuente_quantum_2025} has a natural notion of hypergraph lattices. Could we decompose a Lindbladian as a convex combo of hypergraph lattice Lindbladians?

\section{Local Subalgebras and Observables}
A key consequence of relative entropy contraction is that it controls diamond norm distance, an operational measure of distinguishability, through Pinsker's inequality. The usual relation that $2 \| \rho - \sigma \|_1^2 \leq D(\rho \| \sigma)$ becomes potentially weak for large $n$ and trivial as $n \rightarrow \infty$, since the initial relative entropy could be arbitrarily large. This limitation is physical - existing analyses have shown that the depth required for additive error convergence of random quantum circuits is lower bounded as $\Omega(\log n)$ \cite{dalzell_random_2022, deshpande_tight_2022}. Therefore, even $\Omega(1)$ entropy contraction cannot yield general, constant-depth additive error convergence. Instead, entropy contraction bounds the \textit{average} trace on local subalgebras via the superadditivity of relative entropy on tensor splits and Pinsker's inequality:
\begin{lemma} \label{lem:throwaway}
    Consider channel $\Phi_{\mathrm{Haar}, k}$ acting on the entirety of a $k$-fold bipartite system $A^k B^k$, followed by a partial trace of the $B$ systems. Let $\E_{dep}^A$ denote the completely depolarizing channel on $A^k$. Assuming $|B| \geq 2 k^2$, $\tr_B \circ \Phi_{\mathrm{Haar}, k} \precsucc{\epsilon}{} \E_{dep}^A$ with $\epsilon \leq 2 k^2 / |B|$.
\end{lemma}
\begin{proof}
    First consider the scenario in which $B \cong B_0 \otimes A^{n-1}$ for some $n \in \NN$ and each of the $k$ copies. By Schur-Weyl duality and using the Bell pair transpose trick to apply $\Phi_{\mathrm{Haar},k}$ to both sides,
    \begin{equation}
        D((\Id \otimes \Phi_{\mathrm{Haar},k}) (\mathrm{BP}) \| (\Id \otimes \E_{dep})('')) = O(k \log k)
    \end{equation}
    with no $n$-dependence, where $\Id$ is the identity channel acting on an auxiliary system $A^{'k} B^{'k}$ of the same type as the input, and $\E_{dep}$ is a completely depolarizing channel on system $A^k B^k$. Note that $(\Id \otimes (\E_{dep} \circ \Phi_{\mu,k}))(\mathrm{BP}) = (\Id \otimes \E_{dep})(\mathrm{BP}) = \id / |A A' B B'|^k$ is a product stateon all subsystem divisions. By superadditivity of the relative entropy with respect to tensor products, letting $A_j^k$ denote the $k$-fold system corresponding to the $j$th of $n$ copies of $A^k$,
    \begin{equation}
        \sum_j D((\Id \otimes (\E_{A_j} \circ \Phi_{\mathrm{Haar},k})) (\mathrm{BP}) \| (\Id \otimes \E_{dep})('')) 
            \leq D((\Id \otimes \Phi_{\mathrm{Haar},k}) (\mathrm{BP}) \| (\Id \otimes \E_{dep})('')) \pl.
    \end{equation}
    After applying $\Phi_{\mathrm{Haar}, k}$, the resulting state is invariant under exchanging the $n$ copies of $A^k$. Hence, the individual $A_j$ terms as above are all equal to the average, $D((\Id \otimes \Phi_{\mathrm{Haar},k}) (\mathrm{BP}) \| (\Id \otimes \E_{dep})('')) / n$. To complete the first part of the Lemma, note for any $\delta > 0$ there is an $n$ sufficiently large that the aforementioned is at most $\delta$.

    Let $\Phi_{\mathrm{Haar},k}[BC]$ denote the $k$-fold twirl on $B^k C^k$ (acting as identity on $A$), where $C \cong A^n$ for an $n \in \NN$ to be determined, $\Phi_{\mathrm{Haar},k}[AB] := \Phi_{\mathrm{Haar},k} \otimes \Id^C$, and $\Phi_{\mathrm{Haar},k}[ABC]$ denote the tripartite twirl. By the SHH gluing Lemma \cite[Lemma 3]{schuster_random_2025} (see also \cite{laracuente_approximate_2026}), $\tilde{\Phi}_{\mathrm{Haar},k} \circ \Phi_{\mathrm{Haar},k}$ is a $\kappa$-approximate relative $k$-design with $\kappa \leq 2 k^2 / |B|$ as long as $|B| \geq 2 k^2$. Let $\tilde{E}_A : A^k B^k C^k \rightarrow A^k$ denote the restriction map, and note that
    \begin{equation}
        \Phi_{\mathrm{Haar},k}[BC] \circ  \Phi_{\mathrm{Haar},k}[AB]
            \precsucc{\kappa}{} \Phi_{\mathrm{Haar},k}[ABC] \pl,
    \end{equation}
    where $\Phi_{\mathrm{Haar},k}[BC]$ denotes the Haar twirl on the $BC$ system.
    %Assuming that $\kappa \leq 1/2$, $1/(1 - \kappa) = 1 + \kappa + \kappa \times \kappa/(1+\kappa) \geq 1 + 4 \kappa / 3$. Conversely, $1/(1 + \kappa) \leq 1 - \kappa$. Reversing the order inequality and applying $\tr_{BC}$ to both sides,
    %\begin{equation}
    %        \tr_{BC} \circ \Phi_{\mathrm{Haar},k}[ABC] \precsucc{4 \kappa / 3}{} \tr_{BC} \circ \Phi_{\mathrm{Haar},k}[BC]  \circ \Phi_{\mathrm{Haar},k}[AB] \pl.
    %\end{equation}
    Let $F_\omega : AB \rightarrow ABC$ denote the extension by an arbitrary state $\omega$ on $C^k$.
    \begin{equation}
    \begin{split}
        \tr_{BC} \circ \Phi_{\mathrm{Haar},k}[BC] \circ \Phi_{\mathrm{Haar},k}[AB] \circ F_\omega
            = \tr_{BC} \circ \Phi_{\mathrm{Haar},k}[BC] \circ F_\omega \circ \Phi_{\mathrm{Haar},k}
            = \tr_{B} \circ \Phi_{\mathrm{Haar},k} \pl,
    \end{split}
    \end{equation}
    since the inserted extension and middle twirl act only on systems that will subsequently be discarded. Therefore,
    \begin{equation}
        \tr_{B} \circ \Phi_{\mathrm{Haar},k} \precsucc{\kappa}{} 
            \tr_{BC} \circ \Phi_{\mathrm{Haar},k}[ABC] \circ F_\omega \pl.
    \end{equation}
    Taking $n$ arbitrarily large, $\tr_{BC} \circ \Phi_{\mathrm{Haar},k}[ABC]$ becomes an arbitrarily good approximation to $\E_{dep} \otimes \tr_{BC}$ in relative entropy. By the faithfulness of relative entropy and the Choi-Jamiolkowski isomorphism, these are therefore the same channel. Hence,
    \begin{equation}
        \tr_{B} \circ \Phi_{\mathrm{Haar},k} \precsucc{\kappa}{} 
            \E_{dep} \otimes \tr_{BC} \circ F_\omega = \Id \otimes \E_{dep} \pl,
    \end{equation}
    completing the Lemma.
\end{proof}
\todo{We need a version of this (and underlying it, the gluing Lemma) for fermions.}
\begin{theorem}[Average Subsystem Convergence] \label{prop:avg}
    On $n$ qudits $A_1, \dots, A_n$ of local dimension $q$, for any input state $\rho$,
    \begin{equation*}
    \begin{split}
        \frac{1}{n} \sum_{j \in 1}^n D \left (\rho^{A' A_j} \big \| \rho^{A'} \otimes \id / q \right)
        \leq & \frac{D(\rho \| (\Id \otimes \Phi_{\mathrm{Haar},k})(\rho))}{n (1 - \log_q (24 n k^2) / n)(1 - 1/n)} \pl.
    \end{split}
    \end{equation*}
    where the identity is on an auxiliary system of the same dimension denoted $A'$.
\end{theorem}
\begin{proof}
    Via data processing, relative entropy is non-increasing when discarding systems. Assume we discard a subsystem $B$ of the joint system we call $AB = A_1 \dots A_n$. Let $A'$ denote the auxiliary. By Lemma \ref{lem:throwaway},
    \begin{equation}
        \tr_B \circ \Phi_{\mathrm{Haar},k} \precsucc{\epsilon}{} \E_{dep} \circ \tr_B \pl,
    \end{equation}
    where $\E_{dep}$ is a completely depolarizing map, and $\epsilon \leq 2 k^2 / |B|$. Using \ref{lem:revconv},
    \begin{equation}
        D(\rho \| (\Id \otimes \Phi_{\mathrm{Haar},k})(\rho)) \geq 
            \beta_{\epsilon} D(\tr_{B}(\rho) \| \rho^{A'} \otimes \id/|A|) \pl.
    \end{equation}
    Assume $|A| = q^m$ and without loss of generality consider it to be the first $m$ indexed subsystems. 
    For each $j = 1 \dots m$, using \cite[Proposition 2]{capel_superadditivity_2018}
    \begin{equation}
        D(\rho^{A_j} \| \id / |A_j|) = D(\rho^{A' A_j} \| \rho^{A'} \otimes \id / |A_j|) - I(A' : A_j)_\rho \pl.
    \end{equation}
    Using \cite[Proposition 2]{capel_superadditivity_2018},
    \begin{equation}
    \begin{split}
        & D(\rho^{A'A} \| \rho^{A'} \otimes \id/|A|)
        \\ & = I(A'B' : A)_\rho + D(\tr_{B}(\rho) \| \id/|A|)
        \geq I(A' : A)_\rho + \frac{1}{m} \sum_{j = 1}^m D(\rho^{A_j} \| \id / |A_j|)
        \\ & \geq \frac{1}{m} \sum_{j=1}^m \ D(\rho^{A' A_j} \| \rho^{A'} \otimes \id / |A_j|)
            - \frac{1}{m} \sum_{j=1}^m I(A' : A_j)_\rho + I(A' : A)_\rho \pl.
    \end{split}
    \end{equation}
    The same inequality holds regardless of which subsystems were actually chosen to be the $m$ kept, so by averaging over that choice,
    \begin{equation}
    \begin{split}
        \frac{1}{n} \sum_{j=1}^n \ D(\rho^{A' A_j} \| \rho^{A'} \otimes \id / |A_j|)
            \leq & \frac{1}{m} (1 / \beta_{\epsilon}) D(\rho \| (\Id \otimes \Phi_{\mathrm{Haar},k})(\rho)) 
            \\ & + \EE_A \left [ \frac{1}{m} \sum_{j=1}^m I(A' : A_j)_\rho - I(A' : A)_\rho \right ] \pl,
    \end{split}
    \end{equation}
    where $\EE_A$ denotse the average over the choice of $A$ system, and the $j = 1 \dots m$ sum runs over indices that form $A$. Note that each $I(A : A_j)_\rho \leq I(A' : A)$, upper-bounding the average as well. To fix the choice of $m$ given that $\beta \geq 1 - 12 \epsilon \geq 1 - 24 k^2 / q^{n-m}$, let $m := n - \log_q (24 n k^2)$, so $24 k^2 / q^{n-m} \leq 1/n$.
\end{proof}
\begin{cor}
    Let $\Phi$ be a quantum channel with CMLSI constant $\lambda$ toward fixed point projection $\Phi_{\mathrm{Haar},k}$. Let $\E_1, \dots, \E_{n}$ be a family of restriction maps to disjoint subsystems each of local dimension $q$. Then
    \begin{equation*}
        \frac{1}{n} \sum_{j=1}^{n} \left \| \E_j \circ \Phi - \E_{dep} \right \|_{\Diamond}^2 \leq \frac{ 4 k \exp(-\lambda /2) \log q }{(1 - \log_q (24 n k^2) / n)(1 - 1/n)} = 4 k \exp(-\lambda / 2) \log q  + O \Big ( \frac{1}{n} \Big ) \pl,
    \end{equation*}
    where $\E_{dep}$ replaces its input by $\id / q$, the complete mixture.
\end{cor}
\begin{proof}
    Observe that by the data processing inequality, we may trace out any extra qudits, obtaining an effective system of $n$. Apply Proposition \ref{prop:avg}, bunching qudits into units of size $m$, noting that the maximum, global relative entropy is $2 n k \log q$. Apply Pinsker's inequality following , 
\end{proof}
The next step is to apply:
\begin{enumerate}
    \item Light cone bounds to the initial state.
    \item Translation invariance to the final state.
\end{enumerate}
With these, we should be able to show that \textit{each individual} local observable is close to the Haar-twirled value.
\begin{cor} \label{cor:local}
    Let $\Phi$ be a quantum channel with CMLSI constant $\lambda$ toward fixed point projection $\Phi_{\mathrm{Haar},k}$ on system $AB$. Let $\E_1, \dots, \E_{n}$ be a family of restriction maps to disjoint subsystems each of local dimension $q$ all contained in the light cone of input subsystem $A$, and let $R$ denote an arbitrary reference or auxiliary system. Assume that at the output, the ratio of the largest entropy to the average is at most $\alpha$. Then for each $j = 1 \dots n$ and any input $\rho$,
    \begin{equation}
    \begin{split}
    D \left ((\Id^R \otimes \E_j)(\rho) \big \| \rho^{R} \otimes \id / q \right)
    & \leq \frac{\alpha D(\rho^{RA} \otimes \id /|B| \| (\Id^R \otimes \Phi_{\mathrm{Haar},k})(\rho^{RA} \otimes \id / |B|))}{n (1 - \log_q (24 n k^2) / n)(1 - 1/n)}
    \\ & \leq \frac{2 \alpha \exp(-\lambda) \log |A|}{n (1 - \log_q (24 n k^2) / n)(1 - 1/n)} \pl,
    \end{split}
    \end{equation}
    and
    \begin{equation}
        \left \| \E_j \circ \Phi - \E_{dep} \right \|_{\Diamond}^2
        \leq \frac{4 \alpha \exp(-\lambda / 2) \log |A|}{n (1 - \log_q (24 n k^2) / n)(1 - 1/n)} \pl.
    \end{equation}
\end{cor}

\bibliographystyle{unsrt}
\bibliography{ref}

\end{document}